% §5.1 Load regime
%
% Markdown companion: paper/sections/05_1_load_regime.md.
% Inserted BEFORE Table 1 in the paper per the P10 task spec so the
% reader sees the load-regime context before reading the throughput
% column.

\section{Load regime: what throughput actually measures}
\label{sec:load_regime}

\noindent\textbf{In the experimental regime studied here, the
system-wide task arrival rate equals $|M|/(\text{height} +
\text{width})$ per step, which is 0.781 on
\texttt{random-64-64-10} and 0.394 on
\texttt{warehouse-10-20-10-2-2} at $|M|=100$. Throughput in
Table~\ref{tab:solver_subst_random-64-64-10} saturates at this
arrival cap; the column therefore measures task-arrival-rate
compliance, not planner capacity. Algorithm-discriminating
comparisons in \S\ref{sec:system_health}, \S5.5 (baseline
comparison), and \S5.6 (allocator study) operate at points where
the planner becomes the bottleneck; those are the comparisons in
which throughput is informative.}

\medskip

The simulator's default lifelong task stream
(\texttt{Simulator.\_generate\_task\_stream}) draws inter-arrival
times from an exponential with per-agent mean $H + W$ steps, where
$H$ and $W$ are the map height and width.  The system-wide arrival
rate is therefore $\lambda_{\text{sys}} = |M|/(H+W)$ tasks per
step.  At the headline §5.2 conditions $|M|=100$:

\begin{center}
\small
\begin{tabular}{lrr}
\toprule
Map & $H + W$ & $\lambda_{\text{sys}}$ at $|M|=100$ \\
\midrule
\texttt{random-64-64-10}        & $64 + 64 = 128$  & \textbf{0.781} \\
\texttt{warehouse-10-20-10-2-2} & $84 + 170 = 254$ & \textbf{0.394} \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{CSV emission.}
\texttt{MetricsTracker.finalize} now emits
\texttt{arrival\_rate\_per\_step}
($= \text{total\_released\_tasks}/\text{steps}$) and
\texttt{throughput\_utilization}
($= \text{throughput}/\text{arrival\_rate\_per\_step}$).  Utilization
$\ge 0.95$ is treated as arrival-saturated; the table builder
(\texttt{scripts/evaluation/build\_summary\_tables.py}) appends a
trailing asterisk to such cells so a reader can scan for them
without recomputing the ratio.

\paragraph{When throughput is informative.}
At points where the planner is the bottleneck (\S5.3/\S5.4
FoV/$r_{\text{safe}}$ sweeps, \S5.6 allocator study under a fixed
arrival stream) throughput tracks the planner.  At points where
arrival saturates -- the entire §5.2 Table~1, and high-density
warehouse rows of \S5.5 -- the Util.\ column reads $1.00^{*}$ and
throughput cannot tell two planners apart.  Reading the Util.\
column \emph{plus} the \texttt{deadlock\_count} column from
\S\ref{sec:system_health} catches the worst-case reading error of
treating throughput as a planner-quality metric: LaCAM-blind
deadlocks 100\% of agents in every $|M|=100$ warehouse run but
reports throughput 0.205 vs Ours' 0.396 -- a 1:2 ratio that looks
like a half-as-good planner rather than a catastrophic one.
