% §5.4  System Health Indicators
%
% Generated from logs/paper/baseline_comparison_v2/results.csv via the
% column logic in scripts/evaluation/build_summary_tables.py
% (HEALTH_COLS).  Markdown companion: paper/sections/05_4_system_health.md.

\section{System Health Indicators}
\label{sec:system_health}

\noindent\textbf{Claim.}\ \emph{In configurations where throughput is
task-arrival-limited, the throughput column does not reflect
agent-level progress.  The \texttt{deadlock\_count} column does.}

\medskip

Throughput on the warehouse map at $|M| \ge 100$ tracks the
task-arrival process: every released task is picked up within a few
ticks and completed before the next batch arrives, so throughput
saturates near the arrival rate regardless of how many agents are
functionally stuck.  A subset of agents safety-wait at corridor
junctions for the remainder of the run and contribute zero to
throughput, yet \texttt{completed\_tasks} -- and hence
\texttt{throughput} -- stays near its ceiling because the other
agents pick up the slack.  Without the diagnostics in this section a
reader of Tables~\ref{tab:solver_subst_random-64-64-10}
and~\ref{tab:baselines_100} would see only the saturation regime and
conclude all $|M|$ agents are making progress; in fact a third of
the fleet is silently lost in every high-density warehouse run.

\begin{table}[t]
\centering
\small
\begin{tabular}{rrrrrr}
\toprule
$|M|$ & Throughput & Deadlock count (mean) & Max DL & Frac. with DL & GNP steps \\
\midrule
  50 & 0.204 $\pm$ 0.004 &  5.40 $\pm$ 1.45 &  9 & 100\% &  6.9 $\pm$ 8.6 \\
 100 & 0.396 $\pm$ 0.005 & 16.10 $\pm$ 2.70 & 22 & 100\% &  0.0 $\pm$ 0.0 \\
 150 & 0.596 $\pm$ 0.010 & 20.00 $\pm$ 4.25 & 30 & 100\% &  0.0 $\pm$ 0.0 \\
 200 & 0.795 $\pm$ 0.010 & 17.30 $\pm$ 3.05 & 31 & 100\% &  0.0 $\pm$ 0.0 \\
 250 & 0.992 $\pm$ 0.013 & 20.30 $\pm$ 4.40 & 31 & 100\% &  0.0 $\pm$ 0.0 \\
\bottomrule
\end{tabular}
\caption{Headline cell: \emph{Ours} on warehouse-10-20-10-2-2.
Per-cell mean $\pm$ bootstrap 95\% CI across 10 seeds; max-DL is the
worst-seed deadlock count; Frac.\ with DL is the fraction of seeds
with at least one deadlocked agent.  At $|M|=100$ throughput reads
saturated (0.396), but every seed has at least one deadlocked agent
and an average of 16.1 agents per run cross the deadlock threshold.}
\label{tab:system_health_headline}
\end{table}

\paragraph{Reading the table.}  At $|M| = 100$, throughput shows
0.396 -- the system looks functional.  \texttt{deadlock\_count} says
16.10 agents (mean across seeds) are permanently stuck (16\% of the
fleet) and every seed has at least one deadlocked agent.  As $|M|$
grows, the absolute deadlock count saturates around 20 (about 8\% of
the fleet at $|M|=250$); the fraction of seeds with at least one
deadlocked agent stays at 100\% across every density.  The
worst-seed deadlock count reaches 31 agents simultaneously stuck at
$|M| \ge 200$.

\begin{table}[t]
\centering
\small
\begin{tabular}{lrrr}
\toprule
Method & Deadlock count (mean) & Max DL & Frac. with DL \\
\midrule
Ours (POE-LMAPF) &  16.10 &  22 & 100\% \\
No-Buffer        &  29.30 &  35 & 100\% \\
PIBT2-FR         &  43.00 &  56 & 100\% \\
LaCAM-blind      & 100.00 & 100 & 100\% \\
\bottomrule
\end{tabular}
\caption{Cross-method comparison at $|M|=100$ on
warehouse-10-20-10-2-2.  \emph{LaCAM-blind deadlocks every agent in
every run}: the rigid-follower controller has no recovery path when
the global solution becomes infeasible mid-run.  Reading throughput
alone the method looks half-as-good as Ours; reading
\texttt{deadlock\_count} it is catastrophically broken.  This is the
regression \texttt{deadlock\_count} catches that \texttt{throughput}
does not.}
\label{tab:system_health_methods}
\end{table}

\paragraph{Builder + provenance.}  The four health columns
(\texttt{Deadlock count}, \texttt{Frac.\ runs with DL},
\texttt{Max DL count}, \texttt{GNP steps}) are appended by
\texttt{scripts/evaluation/build\_summary\_tables.py} to every
results table the builder produces; see \texttt{HEALTH\_COLS} in
the builder source.  Source CSVs:
\texttt{logs/paper/baseline\_comparison\_v2/results.csv} for the
tables above and the §5.5 baseline comparison.  The §5.2 sweep CSV
(\texttt{logs/paper/solver\_sensitivity/results.csv}, source for
Table~\ref{tab:solver_subst_random-64-64-10}) predates the
deadlock-count instrumentation and reports em-dashes in the four
health columns; the per-cell footnote under that table points
readers here for the high-density picture.
